---
Title: Automating update of the Christiano, Motto and Rostagno (2014) database for the United States
date: 2016-06-02
Category: Data
Tags: database, model, estimation, R
Slug: cmr14-data
Authors: Thomas Brand, Anatole Cheysson
Summary: Description step by step to build automatic update of the Christiano, Motto and Rostagno (2014) database for the United States.
Download: https://git.nomics.world/macro/cmr14-data
output: html_document
---

```{r, message=FALSE, warning=FALSE, echo=FALSE, results='hide'}
if (!"pacman" %in% installed.packages()[,"Package"]) install.packages("pacman", repos='http://cran.r-project.org')
pacman::p_load(tidyverse,rdbnomics,magrittr,lubridate,knitr,zoo)

pacman::p_load_gh("thomasbrand/FredR")
api.key <- "820ed4f91f76c8fe560514930425118c"
fred <- FredR(api.key)

opts_chunk$set(message=FALSE, warning=FALSE, cache=FALSE, fig.align="center")

theme <-  theme_bw()+ theme(strip.background=element_blank(),
                            strip.text=element_text(size=15),
                            title=element_text(size=16),
                            panel.border=element_blank(),
                            panel.grid.major=element_line(size=1),
                            legend.key=element_rect(colour="white"),
                            legend.position="bottom",
                            legend.text=element_text(size=10),
                            axis.text=element_text(size=10),
                            plot.title=element_text(hjust=0.5))
blueObsMacro <- "#0D5BA4"
```


Our purpose is to write a program to automatically update a database similar to the one used in the bayesian estimation of the DSGE model developed in [@Chri14a] (hereafter CMR) for the United States.

We need 12 series for the estimation of the CMR model:

1. GDP
2. Deflator of GDP
3. Consumption
4. Investment
5. Deflator of investment
6. Wages
7. Hours worked
8. Loans to non financial corporations
9. Short-term interest rate
10. Net worth
11. Credit spread 
12. Term premium 

To those 12 series we added 2 others:

13. Loans to households and non-profit institutions serving households
14. House prices 

To create this database, many sources are used (BEA, BIS, BLS, OECD). Mostly, we take data directly from <a href="https://db.nomics.world/" target="_blank">DBnomics</a>, through the <a href="https://cran.r-project.org/web/packages/rdbnomics/index.html" target="_blank">rdbnomics</a> package. For Wilshire 5000 Total Market Index and Moody's Seasoned Baa Corporate Bond Yield, we use data from <a href="https://fred.stlouisfed.org/" target="_blank">FRED</a> (from Saint-Louis Fed) using the plugin function <a href="http://www.jankocizel.com/FredR/" target="_blank">FredR</a> from Janko Cizel, which is <a href="https://github.com/thomasbrand/fredR" target="_blank">updated</a>. All the following code is written in R, thanks to the [@Rct16] and the [@RStu16].

# Raw data from BEA, BIS, BLS and OECD

We use DBnomics to retrieve data from BEA, BIS, BLS and OECD.

```{r}
df <- rdb(ids=c("BEA/NIPA-T10106/A191RX-Q",
                "BEA/NIPA-T10109/A191RD-Q",
                "BEA/NIPA-T10106/A006RX-Q",
                "BEA/NIPA-T10109/A006RD-Q",
                "BIS/total_credit/Q.US.N.A.M.XDC.A",
                "BIS/total_credit/Q.US.H.A.M.XDC.A",
                "BIS/selected_pp/Q.US.N.628",
                "BLS/pr/PRS85006033",
                "BLS/pr/PRS85006103",
                "OECD/MEI/USA.IRLTLT01.ST.Q",
                "OECD/MEI/USA.LFWA64TT.STSA.Q")) %>% 
  mutate(series_name=case_when(str_detect(series_code,"RD-") ~ paste("Deflator,",series_name),
                               str_detect(series_code,"RX-") ~ paste("Real,",series_name),
                               str_detect(series_code,"Q.US.N.A.M.XDC.A") ~ paste("Loans to non-financial corporations,",series_name),
                               str_detect(series_code,"Q.US.H.A.M.XDC.A") ~ paste("Loans to households and NPISHs,",series_name),
                               str_detect(series_code,"Q.US.N.628") ~ paste("Property prices,",series_name),
                               TRUE ~ series_name)) %>% 
  select(var_name=series_name,
         var_code=series_code,
         value,
         period)

df %<>% 
  mutate(var_code=
           case_when(var_code=="A191RX-Q" ~ "gdp",
                     var_code=="A006RX-Q" ~ "inves",
                     var_code=="A191RD-Q" ~ "defgdp",
                     var_code=="A006RD-Q" ~ "definves",
                     var_code=="Q.US.H.A.M.XDC.A" ~ "loans_hh",
                     var_code=="Q.US.N.A.M.XDC.A" ~ "loans_nfc",
                     var_code=="Q.US.N.628" ~ "houseprice",
                     var_code=="PRS85006033" ~ "hours",
                     var_code=="PRS85006103" ~ "wage",
                     var_code=="USA.LFWA64TT.STSA.Q" ~ "pop",
                     var_code=="USA.IRLTLT01.ST.Q" ~ "longrate"))
```

We create a quarterly series of short-term interest rate taking the mean of the monthly series.

```{r}
shortrate <- 
  rdb("FED","H15",mask="129.FF.O") %>% 
  mutate(period=paste(year(period),quarter(period),sep="-")) %>% 
  group_by(period) %>% 
  summarise(value=mean(value)) %>% 
  mutate(var_code="shortrate",
         var_name="Monthly – Federal funds – Overnight",
         period=yq(period))
```


# Special case of consumption

Private consumption is a special case : the series of consumption in level provided by the Bureau of Economic Analysis do not exist before 2002, we need to use the series of growth rates available before 2002 to deduce past consumption levels. As in [@Chri14a], aggregate consumption is the sum of Non Durable Goods and Services. Durable Goods are associated with Investment. The following figure shows the three deduced series of consumption.

```{r, fig.height=3.5, fig.width=8}
conso_level <- 
  rdb(ids=c("BEA/NIPA-T20306/DDURRX-Q",
            "BEA/NIPA-T20306/DNDGRX-Q",
            "BEA/NIPA-T20306/DSERRX-Q")) %>% 
  select(period,
         value,
         var_name=concept)

conso_rate <- 
  rdb(ids=c("BEA/NIPA-T20301/DDURRL-Q",
            "BEA/NIPA-T20301/DNDGRL-Q",
            "BEA/NIPA-T20301/DSERRL-Q")) %>% 
  select(period,
         value,
         var_name=concept)

conso_level_02 <-
  conso_level %>% 
  filter(period=="2002-01-01")

conso <-
  conso_rate %>% 
  filter(period <= "2002-01-01") %>% 
  full_join(conso_level_02,by="var_name") %>% 
  group_by(var_name) %>% 
  arrange(desc(period.x)) %>% 
  mutate(value = value.y / lag(cumprod((1 + value.x/100)^(1/4)))) %>%
  ungroup() %>% 
  transmute(period=period.x,
            var_name,
            value) %>% 
  na.omit() %>% 
  bind_rows(conso_level) %>% 
  filter(period >= "1980-01-01")

ggplot(conso,aes(period,value))+
  geom_line(colour=blueObsMacro)+
  facet_wrap(~var_name,ncol=3,scales = "free_y")+
  scale_x_date(expand = c(0.01,0.01)) +
  theme + xlab(NULL) + ylab(NULL)+
  ggtitle("Real Personal Consumption Expenditures")

conso %<>% 
  mutate(
    var_code=case_when(
      var_name=="durable-goods" ~ "conso_d",
      var_name=="nondurable-goods" ~ "conso_nd",
      var_name=="services" ~ "conso_s"
    ),
    var_name=paste("Real Personal Consumption Expenditures,",var_name))
```


# Financial data from FRED

```{r}
fred_data <- 
  bind_rows(data.frame(fred$series.observations(series_id = 'BAA',
                                                frequency="q", 
                                                aggregation='avg'), 
                       var_name="Moody's Seasoned Baa Corporate Bond Yield",
                       var_code="riskrate"), 
            data.frame(fred$series.observations(series_id = 'WILL5000IND', 
                                                frequency="q", 
                                                aggregation='avg'), 
                       var_name="Wilshire 5000 Total Market Index",
                       var_code="networth")) %>% 
  transmute(period=ymd(date),
            value=as.numeric(value),
            var_code,var_name)
```

# Final database and normalization

```{r}
rawdata <- 
  bind_rows(conso,df,shortrate,fred_data) %>%
  filter(year(period) >= 1980)
```

```{r}
var_names <- unique(rawdata$var_name)
var_names <- gsub("Expenditures,.*","",var_names) %>% unique()
```

We can check the last date available for each variable.
```{r}
maxDate <- 
  rawdata %>% 
  filter(var_code!="houseprice") %>% 
  group_by(var_code) %>% 
  summarize(maxdate=max(period)) %>% 
  arrange(maxdate)
kable(maxDate)
```

```{r}
minmaxDate <- min(maxDate$maxdate)
rawdata %<>% filter(period <= minmaxDate) %>% select(-var_name)
rawdata %>%
  spread(var_code,value) %>%
  write.csv(file = "US_CMR_rawdata.csv", row.names= FALSE)
```

So we filter the database until `r as.yearqtr(minmaxDate)`. Raw data can be downloaded directly <a href="http://shiny.cepremap.fr/data/US_CMR_rawdata.csv" target="_blank">here</a>.

Then data are normalized by capita and price if needed. Eventually we have the 14 series : the 12 series similar to CMR plus loans to households and house price series.

```{r}
US_CMR_data <- 
  rawdata %>%
  spread(var_code,value) %>% 
  transmute(period,
            gdp_rpc=1e+6*gdp/(1000*pop)/4,
            conso_rpc=1e+6*(conso_nd+conso_s)/(1000*pop)/4,
            inves_rpc=1e+6*(inves+conso_d)/(1000*pop)/4,
            defgdp = defgdp,
            wage_rph=wage/defgdp,
            hours_pc=1e+9*hours/(1000*pop),
            pinves_defl=definves/defgdp,
            loans_nfc_rpc=1e+9*loans_nfc/(1000*pop)/defgdp,
            loans_hh_rpc=1e+9*loans_hh/(1000*pop)/defgdp,
            houseprice_defl=houseprice/defgdp,
            networth_rpc=1e+6*networth/(1000*pop)/defgdp,
            re=shortrate/100,
            slope=(longrate - shortrate)/100,
            creditspread=(riskrate - longrate)/100)

US_CMR_data %>%
  mutate(period=gsub(" ","",as.yearqtr(as.Date(period)))) %>%
  write.csv("US_CMR_data.csv", row.names=FALSE)
```

```{r echo=FALSE}
listVar <- list("Real GDP per capita"                      = "gdp_rpc",
                "Real consumption per capita"              = "conso_rpc",
                "Real investment per capita"               = "inves_rpc",
                "Real credit to NFC per capita"            = "loans_nfc_rpc",
                "Real credit to HH per capita"             = "loans_hh_rpc",
                "Real house price"                         = "houseprice_defl",
                "Real net worth per capita"                = "networth_rpc",
                "Real price of investment"                 = "pinves_defl",
                "Real wage per capita"                     = "wage_rph",
                "GDP deflator"                             = "defgdp",
                "Hours worked per capita"                  = "hours_pc",
                "Short-term interest rate (APR)"           = "re",
                "Credit spread (APP)"                      = "creditspread",
                "Term premium (APP)"                       = "slope")
```

```{r, fig.height=8.5, fig.width=8}
plot_US_CMR_data <- 
  US_CMR_data %>% 
  gather(var, value, - period) %>%
  mutate(var = as.factor(var))
levels(plot_US_CMR_data$var)<-listVar

ggplot(plot_US_CMR_data,aes(period,value))+
  geom_line(colour=blueObsMacro)+
  facet_wrap(~var,ncol=3,scales = "free_y")+
  scale_x_date(expand = c(0.01,0.01)) +
  theme + xlab(NULL) + ylab(NULL)+
  theme(strip.text=element_text(size=10),
        axis.text=element_text(size=9))+
  ggtitle("CMR data for the US")
```

You can also download ready-to-use (normalized) data for the estimation <a href="http://shiny.cepremap.fr/data/US_CMR_data.csv" target="_blank">here</a>.
